Precise Data Flow Analysis\\in the Presence of Correlated Method Calls

### 2015-03-12 sas2015/content/example.tex
fiddle with figures
move figures around to save half a page
proofreading pass
minor changes

Figure~\ref{list:ccexample} shows a small Java program.
%that we will use to illustrate
%how our technique improves the precision of a simple taint analysis.
The program
declares a class \code{A} with a subclass \code{B},  where \code{A} defines methods 
\code{foo()} and \code{bar()} that are overridden in \code{B}.  
%To keep the 
%example small and self-contained, it is 
We assume  that secret values are created by 
an unspecified function \code{secret()}, which is called in \code{A.foo()} on 
line~\ref{line:Afoo}. 
%Furthermore, it is assumed that 
Any write to standard output 
%via calls \code{System.out.println()} such as the one in method \code{B.bar()} 
%on line~\ref{line:Bbar} 
is assumed to be a sink (e.g. the call to \code{System.out.println()} in code{B.bar()}.
%a security-sensitive operation.
Depending on the number of 
arguments passed to the program, the \code{main()} method of the example program creates 
either an \code{A}-object or a \code{B}-object%
%on line~\ref{line:createObject}
. The 
program then calls \code{foo()} on this object
on line~\ref{line:callfoo}, which is 
followed by a call to \code{bar()} on the same object%
%on line~\ref{line:callbar}
.

@@

As we discussed previously, the calls to \code{foo()} and \code{bar()} may dispatch
to the implementations in classes \code{A} and \code{B}, 
because the receiver variable \code{a} may be bound to
 objects of type \code{A} or \code{B} at run time. 
However, the methods \code{foo()} and \code{bar()} are invoked
on \textit{the same object}. Thus the behaviours of the method calls
are \textit{correlated}: if the call to \code{foo()} dispatches to \code{A.foo()},
then the call to \code{bar()} must dispatch to \code{A.bar()}, and analogously
for \code{B.foo()} and \code{B.bar()}.
Consequently, paths such as the one highlighted in
Figure~\ref{fig:examplesupergraph} where the calls dispatch to
\code{A.foo()} and \code{B.bar()}
are infeasible.

@@

In these programs, on average, 3\% of all call sites $C$ are polymorphic call sites $C_P$.
%\mtodo{Should we explain in a footnote why the number of polymorphic call sites is so low?}
%\otodo{We don't have an explanation.}
Out of these polymorphic call sites, a significant fraction (39\%) are correlated 
call sites $C^\Subset$. We also see that, on average,  each correlated-call receiver is involved in three correlated calls.

>>>>

Figure~\ref{list:ccexample} shows a small Java program.
%that we will use to illustrate
%how our technique improves the precision of a simple taint analysis.
The program
declares a class \code{A} with a subclass \code{B},  where \code{A} defines methods 
\code{foo()} and \code{bar()} that are overridden in \code{B}.  
%To keep the 
%example small and self-contained, it is 
We assume  that secret values are created by 
an unspecified function \code{secret()}, which is called in \code{A.foo()} on 
line~\ref{line:Afoo}. 
%Furthermore, it is assumed that 
Any write to standard output 
%via calls \code{System.out.println()} such as the one in method \code{B.bar()} 
%on line~\ref{line:Bbar} 
is assumed to be a sink (e.g., the call to \code{System.out.println()} in \code{B.bar()}).
%a security-sensitive operation.
Depending on the number of 
arguments passed to the program, the \code{main()} method of the example program creates 
either an \code{A}-object or a \code{B}-object%
%on line~\ref{line:createObject}
. The 
program then calls \code{foo()} on this object
on line~\ref{line:callfoo}, which is 
followed by a call to \code{bar()} on the same object%
%on line~\ref{line:callbar}
.

@@

As we discussed previously, the calls to \code{foo()} and \code{bar()} may dispatch
to the implementations in classes \code{A} and \code{B}, 
because the receiver variable \code{a} may be bound to
 objects of type \code{A} or \code{B} at run time. 
However, the methods \code{foo()} and \code{bar()} are invoked
on \textit{the same object}. Thus the behaviours of the method calls
are \textit{correlated}: if the call to \code{foo()} dispatches to \code{A.foo()},
then the call to \code{bar()} must dispatch to \code{A.bar()}, and analogously
for \code{B.foo()} and \code{B.bar()}.
Consequently, paths such as the one shown in bold in
Figure~\ref{fig:examplesupergraph} where the calls dispatch to
\code{A.foo()} and \code{B.bar()}
are infeasible.

@@

In these programs, on average, 3\% of all call sites $C$ are polymorphic call sites $C_P$.
%\mtodo{Should we explain in a footnote why the number of polymorphic call sites is so low?}
%\otodo{We don't have an explanation.}
Out of these polymorphic call sites, a significant fraction (39\%) are correlated 
call sites $C^\Subset$. We also see that, on average,  each correlated-call receiver is involved in approximately
three correlated calls.

### 2015-03-12 sas2015/content/intro.tex
minor changes

%Since a CFG is an over-approximation of the possible flows 
A control-flow graph (CFG) is an over-approximation of the possible flows 
of control in concrete executions of a program. It may contain \textit{infeasible} 
paths that cannot occur at runtime. The precision of a data-flow analysis algorithm
depends on its ability to detect and disregard such infeasible paths.
The \textit{Interprocedural Finite Distributive Subset} (IFDS) algorithm~\cite{reps1995precise}
is one general data-flow analysis algorithm that 
%by  Reps, Horwitz, and Sagiv
%is a general data-flow analysis
%algorithm
%for computing solutions to finite distributive data-flow problems such 
%that can be instantiated to solve problems such
%as reaching definitions, available expressions, and taint analysis.
%A distinguishing 
%characteristic of IFDS is that it
avoids infeasible interprocedural paths in which calls 
and returns to/from functions are not properly matched. %Sagiv, Reps, and Horwitz
%also presented
The \textit{Interprocedural Distributive Environment} (IDE) algorithm~\cite{sagiv1996precise}
has the same property, but
%that similarly only considers properly matched call/return
%edges, but that 
supports a broader range of data-flow problems.
%by expanding the 
%domain of flow functions to \textit{environments} that go beyond the data-flow facts 
%considered by IFDS.

This paper presents an approach to data-flow analysis that avoids a type
of infeasible path that arises in object-oriented programs when two or more
methods are dynamically dispatched on the same receiver object.
%\mtodo{Should we define the term ``receiver''?}
%In such cases,
If the
method calls are polymorphic
(i.e. the method invoked depends on the run-time type of the receiver),
%(i.e., if they dispatch to different method definitions 
%depending on the type of the receiver expression at run time),
then their dispatch 
behaviours are correlated, and some of the paths between them are infeasible. A recent paper~\cite{DBLP:journals/scp/Tip15} made this observation
but did not present any concrete algorithm to take advantage of it.
%, and we are not aware 
%of any existing data-flow analysis that is capable of avoiding infeasible paths that arise 
%in the presence of correlated method calls.
% (see Section~\ref{sec:Related} for a discussion
%of related work).

@@

The remainder of this paper is organized as follows.
%
Section~\ref{sec:MotivatingExample} presents a motivating example.
%
Section~\ref{sec:bg} reviews the IFDS and IDE algorithms.
%
Section~\ref{sec:cca} presents the correlated-calls transformation
and sketches a proof of its correctness%
\footnote{
  Detailed proofs of our lemmas and theorems can be found in
  the appendix.
}.
%
Related work is discussed in Section~\ref{sec:Related}.
%
Finally, conclusions and directions for future work are presented in Section~\ref{sec:Conclusions}.

>>>>

%Since a CFG is an over-approximation of the possible flows 
A control-flow graph (CFG) is an over-approximation of the possible flows 
of control in concrete executions of a program. It may contain \textit{infeasible} 
paths that cannot occur at runtime. The precision of a data-flow analysis algorithm
depends on its ability to detect and disregard such infeasible paths.
The \textit{Interprocedural Finite Distributive Subset} (IFDS) algorithm~\cite{reps1995precise}
is a general data-flow analysis algorithm that 
%by  Reps, Horwitz, and Sagiv
%is a general data-flow analysis
%algorithm
%for computing solutions to finite distributive data-flow problems such 
%that can be instantiated to solve problems such
%as reaching definitions, available expressions, and taint analysis.
%A distinguishing 
%characteristic of IFDS is that it
avoids infeasible interprocedural paths in which calls 
and returns to/from functions are not properly matched. %Sagiv, Reps, and Horwitz
%also presented
The \textit{Interprocedural Distributive Environment} (IDE) algorithm~\cite{sagiv1996precise}
has the same property, but
%that similarly only considers properly matched call/return
%edges, but that 
supports a broader range of data-flow problems.
%by expanding the 
%domain of flow functions to \textit{environments} that go beyond the data-flow facts 
%considered by IFDS.

This paper presents an approach to data-flow analysis that avoids a type
of infeasible path that arises in object-oriented programs when two or more
methods are dynamically dispatched on the same receiver object.
%\mtodo{Should we define the term ``receiver''?}
%In such cases,
If the
method calls are polymorphic
(i.e., the method invoked depends on the run-time type of the receiver),
%(i.e., if they dispatch to different method definitions 
%depending on the type of the receiver expression at run time),
then their dispatch 
behaviours are correlated, and some of the paths between them are infeasible. A recent paper~\cite{DBLP:journals/scp/Tip15} made this observation
but did not present any concrete algorithm to take advantage of it.
%, and we are not aware 
%of any existing data-flow analysis that is capable of avoiding infeasible paths that arise 
%in the presence of correlated method calls.
% (see Section~\ref{sec:Related} for a discussion
%of related work).

@@

The remainder of this paper is organized as follows.
%
Section~\ref{sec:MotivatingExample} presents a motivating example.
%
Section~\ref{sec:bg} reviews the IFDS and IDE algorithms.
%
Section~\ref{sec:cca} presents the correlated-calls transformation
and sketches a proof of its correctness%
\footnote{
  Detailed proofs of our lemmas and theorems can be found in
  the appendix.
}.
%
Our implementation is discussed in Section~\ref{sec:Evaluation}.
%
Related work is discussed in Section~\ref{sec:Related}.
%
Finally, conclusions and directions for future work are presented in Section~\ref{sec:Conclusions}.

### 2015-03-12 sas2015/content/related.tex
minor changes

The IFDS algorithm is implemented in two popular static-analysis frameworks for Java bytecode, the T.J.~Watson Libraries for~Analysis (WALA)~\cite{fink2012wala} and Heros~\cite{bodden2012inter}.

@@

Sridharan et~al.'s improvements to the precision of pointer analysis
for Ja\-va\-Script programs~\cite{DBLP:conf/ecoop/SridharanDCST12} also
eliminate infeasible paths. Instead of infeasible paths
between dynamically dispatched method calls, their approach eliminates
infeasible paths between reads and writes of different members
%(the JavaScript analogue of fields) 
of an object. 
%Unlike in more static languages,
%property names in JavaScript are first class values (strings). When a loop indexed by
%property names updates the properties of objects, the loop body is extracted into a function
%and analyzed context sensitively with a separate context for each property name.
The approach differs from ours in that it targets points-to analysis rather
than IFDS analyses, in that it targets infeasible paths due to different
property names rather than different dynamically dispatched methods,
and in that it employs context sensitivity to improve precision.

>>>>

%The IFDS algorithm is implemented in two popular static-analysis frameworks for Java bytecode, the T.J.~Watson Libraries for~Analysis (WALA)~\cite{fink2012wala} and Heros~\cite{bodden2012inter}.

@@

Sridharan et~al.'s improvements to the precision of pointer analysis
for Ja\-va\-Script programs~\cite{DBLP:conf/ecoop/SridharanDCST12} also
eliminate infeasible paths. Instead of infeasible paths
between dynamically dispatched method calls, their approach eliminates
infeasible paths between reads and writes of different properties
%(the JavaScript analogue of fields) 
of an object. 
%Unlike in more static languages,
%property names in JavaScript are first class values (strings). When a loop indexed by
%property names updates the properties of objects, the loop body is extracted into a function
%and analyzed context sensitively with a separate context for each property name.
The approach differs from ours in that it targets points-to analysis rather
than IFDS analyses, in that it targets infeasible paths due to different
property names rather than different dynamically dispatched methods,
and in that it employs context sensitivity to improve precision.
